---
title: "Submission 1"
author: "Diana Dima"
date: "7/23/2021"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
---
title: "Mushrooms"
output:
  html_document:
    df_print: paged
  github_document:
    html_preview: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#loading libraries

library(caret)
library(randomForest)  
library(caTools)  
library(rpart)  
library(rpart.plot)  
library(ggplot2)  
library(readr)
library(corrplot)
library(tidyverse)
library(ROCR)
library(factoextra)

setwd("~/Desktop/final")
mushrooms <- read.csv(file = 'mushrooms.csv')
head(mushrooms)
summary(mushrooms)
str(mushrooms)

#Transforming variable observations from character -> factor for bar plots
mushrooms[sapply(mushrooms, is.factor)] <- lapply(mushrooms[sapply(mushrooms, is.factor)], as.character) 
str(mushrooms)

#EDA - bar charts for each attribute

#CLASS
ggplot(mushrooms, aes(`class`, fill = `class`)) + geom_bar() + labs(title =
                    "Class Stacked Bar Chart", x = "Cap Shape" , y = "Count") +                     scale_fill_hue(direction = -1)
                  
#CAP SHAPE
ggplot(mushrooms, aes(`cap.shape`, fill = `class`)) + geom_bar() + labs(title =
                    "Cap Shape Stacked Bar Chart", x = "Cap Shape" , y = "Count") +             scale_fill_hue(direction = -1)

#CAP SURFACE
ggplot(mushrooms, aes(`cap.surface`, fill = `class`)) + geom_bar() + labs(title =
                     "Cap Surface Stacked Bar Chart", x = "Cap Surface" , y = "Count") +             scale_fill_hue(direction = -1)

#CAP COLOR
ggplot(mushrooms, aes(`cap.color`, fill = `class`)) + geom_bar() + labs(title =
                   "Cap Color Stacked Bar Chart", x = "Cap Color" , y = "Count") + 
                   scale_fill_hue(direction = -1)

#BRUISES
ggplot(mushrooms, aes(`bruises`, fill = `class`)) + geom_bar() + labs(title =
                     "Bruises Stacked Bar Chart", x = "Bruises" , y = "Count") +
                     scale_fill_hue(direction = -1)

#ODOR
ggplot(mushrooms, aes(`odor`, fill = `class`)) + geom_bar() + labs(title =
                     "Odor Stacked Bar Chart", x = "Odor" , y = "Count") +
                     scale_fill_hue(direction = -1)
                 
#GILL ATTACHMENT
ggplot(mushrooms, aes(`gill.attachment`, fill = `class`)) + geom_bar() + labs(title =
                     "Gill Attachment Stacked Bar Chart", x = "Gill Attachment" , y = "Count")+
                     scale_fill_hue(direction = -1)


#GILL SPACING
ggplot(mushrooms, aes(`gill.spacing`, fill = `class`)) + geom_bar() + labs(title =
                     "Gill Spacing Stacked Bar Chart", x = "Gill Spacing" , y = "Count") +
                     scale_fill_hue(direction = -1)


#GILL SIZE
ggplot(mushrooms, aes(`gill.size`, fill = `class`)) + geom_bar() + labs(title =
                    "Gill Size Stacked Bar Chart", x = "Gill Size" , y = "Count") +
                     scale_fill_hue(direction = -1)


#GILL COLOR
ggplot(mushrooms, aes(`gill.color`, fill = `class`)) + geom_bar() + labs(title =
                     "Gill Color Stacked Bar Chart", x = "Gill Color" , y = "Count")+
                     scale_fill_hue(direction = -1)


#STALK SHAPE
ggplot(mushrooms, aes(`stalk.shape`, fill = `class`)) + geom_bar() + labs(title =
                    "Stalk Shape Stacked Bar Chart", x = "Stalk Shape" , y = "Count") +
                     scale_fill_hue(direction = -1)


#STALK ROOT
ggplot(mushrooms, aes(`stalk.root`, fill = `class`)) + geom_bar() + labs(title =
                     "Stalk Root Stacked Bar Chart", x = "Stalk Root" , y = "Count")+
                     scale_fill_hue(direction = -1)



#STALK SURFACE ABOVE RING
ggplot(mushrooms, aes(`stalk.surface.above.ring`, fill = `class`)) + geom_bar() + labs(title =
 "Stalk Surface Above Ring Stacked Bar Chart", x = "Stalk Surface Above Ring" , y = "Count")+
                     scale_fill_hue(direction = -1)


#STALK SURFACE BELOW RING 
ggplot(mushrooms, aes(`stalk.surface.below.ring`, fill = `class`)) + geom_bar() + labs(title =
                     "Stalk Surface Below Ring Stacked Bar Chart", x = "Stalk Surface Below Ring" , y = "Count") + scale_fill_hue(direction = -1)


#STALK COLOR ABOVE RING
ggplot(mushrooms, aes(`stalk.color.above.ring`, fill = `class`)) + geom_bar() + labs(title =
                     "Stalk Color Above Ring Stacked Bar Chart", x = "Stalk Color Above Ring" , y = "Count")+ scale_fill_hue(direction = -1)


#STALK COLOR BELOW RING
ggplot(mushrooms, aes(`stalk.color.below.ring`, fill = `class`)) + geom_bar() + labs(title =
                  "Stalk Color Below Ring Stacked Bar Chart", x = "Stalk Color Below Ring" , y = "Count")+ scale_fill_hue(direction = -1)

#VEIL TYPE
ggplot(mushrooms, aes(`veil.type`, fill = `class`)) + geom_bar() + labs(title =
                     "Veil Type Stacked Bar Chart", x = "Veil Type" , y = "Count")+ scale_fill_hue(direction = -1)

#VEIL COLOR
ggplot(mushrooms, aes(`veil.color`, fill = `class`)) + geom_bar() + labs(title =
                     "Veil Color Stacked Bar Chart", x = "Veil Color" , y = "Count")+ scale_fill_hue(direction = -1)


#RING NUMBER
ggplot(mushrooms, aes(`ring.number`, fill = `class`)) + geom_bar() + labs(title =
                    "Ring Number Stacked Bar Chart", x = "Ring Number" , y = "Count")+ scale_fill_hue(direction = -1)


#RING TYPE
ggplot(mushrooms, aes(`ring.type`, fill = `class`)) + geom_bar() + labs(title =
                   "Ring Type Stacked Bar Chart", x = "Ring Type" , y = "Count")+ scale_fill_hue(direction = -1)


#SPORE PRINT COLOR
ggplot(mushrooms, aes(`spore.print.color`, fill = `class`)) + geom_bar() + labs(title =
                   "Spore Print Color Stacked Bar Chart", x = "Spore Print Color" , y = "Count")+ scale_fill_hue(direction = -1)


#POPULATION
ggplot(mushrooms, aes(`population`, fill = `class`)) + geom_bar() + labs(title =                     "Population Stacked Bar Chart", x = "Population" , y = "Count") + scale_fill_hue(direction = -1)

#HABITAT
ggplot(mushrooms, aes(`habitat`, fill = `class`)) + geom_bar() + labs(title =
                     "Habitat Stacked Bar Chart", x = "Habitat" , y = "Count")+ scale_fill_hue(direction = -1)


#Correlation Matrix

library(corrplot)
library(RColorBrewer)

#Changing variables from factor to numeric for correlation matrix

mushrooms[sapply(mushrooms, is.factor)] <- lapply(mushrooms[sapply(mushrooms, is.factor)], as.numeric) 
str(mushrooms)


head(mushrooms)

res <- cor(mushrooms)
round(res, 2)

install.packages("Hmisc")
library(Hmisc)      

res2 <- rcorr(as.matrix(mushrooms))
res2

#Extracting the correlation coefficients

res2$r


#Making a corrplot

library(corrplot)
corrplot(res, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
         
table_cor_mushrooms <- cor(mushrooms)     
corrplot(table_cor_mushrooms, method="ellipse")

corrplot(table_cor_mushrooms, method="number")


#DATA CLEANING

#Checking  how many variables per class

mushrooms[sapply(mushrooms, is.numeric)] <- lapply(mushrooms[sapply(mushrooms, is.numeric)], as.factor) 

str(mushrooms) 

mushroomClasses<-cbind.data.frame(Var=names(mushrooms), Total_Class=sapply(mushrooms,function(x){as.numeric(length(levels(x)))}))
mushroomClasses


#removing veil type attribute with only one level, and is therefore not deterministic

mushrooms$veil.type <- NULL

#Updating corrplot without veil type attribute

mushrooms[sapply(mushrooms, is.factor)] <- lapply(mushrooms[sapply(mushrooms, is.factor)], as.numeric) 

table_cor_mushrooms <- cor(mushrooms)     
corrplot(table_cor_mushrooms, method="ellipse")


#__________________________________________________

#MODELING

str(mushrooms)

#Changing variables from numeric back to factor

mushrooms[sapply(mushrooms, is.numeric)] <- lapply(mushrooms[sapply(mushrooms, is.numeric)], as.factor) 

str(mushrooms)


#MODELING  - RANDOM FOREST - FEATURE IMPORTANCE

set.seed(101) 

classRandomFor<- randomForest(class ~ ., data=mushrooms, ntree=1000, keep.forest=FALSE, importance=TRUE)

varImpPlot(classRandomFor, main = "Variable Importance - Mushroom Class")

# Odor is the feature that most affects class, followed by Spore Print Color and Gill Color 
str(mushrooms)

mushrooms[sapply(mushrooms, is.character)] <- lapply(mushrooms[sapply(mushrooms, is.character)], as.factor) 

mushrooms <- data.frame(lapply(mushrooms, as.factor))



#Checking Odor relationship with Class 
odor_var <- table(mushrooms$class, mushrooms$odor)
odor_var

#Checking Spore Print Color relationship with class
sporCol <- table(mushrooms$class, mushrooms$spore.print.color)
sporCol


#TRAIN, TEST & SPLIT

sample = sample.split(mushrooms$class, SplitRatio = 0.7)

trainMush <- mushrooms[sample, ]
testMush <- mushrooms[-sample, ]

y_train <- trainMush$class
y_test <- testMush$class


#EDIBLE = 1, POISONOUS = 2

y <- mushrooms$class
head(y)

class(y)
class(mushrooms$class)

train <- trainMush
head(train)



#RANDOM FOREST CLASSIFIER

set.seed(101)

rfModel <- randomForest(y~., data=mushrooms, 
                             ntree=100, 
                             mtry=2, 
                             importance=TRUE,
                             proximity=TRUE)
rfModel

#Based on the confusion matrix, the Random Forest predicts no poisonous mushrooms as edible 
#and no edible mushrooms as poisonous - Random forest predicts with 100% accuracy

#Checking variable importance again

options(repr.plot.height=10, repr.plot.width=10)
varImpPlot(rfModel)

head(importance(rfModel))

#Predicting on the Model with Test data

test_predicted<-predict(rfModel,testMush)

dataf<-data.frame(Orig=testMush,Pred=test_predicted)
dataf


y_predicted<-predict(rfModel,testMush)

df1<-data.frame(Orig=y_test,Pred=y_predicted)

confusionMatrix(table(df1$Orig,df1$Pred)) 

##Predicted on test with 100% accuracy


#Random Forest

df <- data.frame(sapply(mushrooms, function (x) as.numeric(as.factor(x))))
df$class <- df$class - 1
nrows <- nrow(df)
set.seed(314)
model <- sample(1:nrow(df), 0.7 * nrows)

#splitting data

train1 = df[model,]
test2 =   df[-model,]
mNames <- names(train1)


rfModel2 <- as.formula(paste("class ~", paste(mNames[!mNames %in% "class"], collapse = " + ")))
train.rf <- randomForest(rfModel2,train1,ntree=100,importance=T)



#Bringing test data to model

test2$predicted <- round(predict(train.rf ,test2),0)
testset_rf <- test2$predicted;

rfConfusionMatrix <- function(test2, sTitle) {
    tst <- data.frame(test2$predicted, test2$class)
    opts <- c("Predicted", "True")
    names(tst) <- opts
    ma <- plyr::count(tst)
    ma[opts][ma[opts]==0] <- "P"
    ma[opts][ma[opts]==1] <- "E"
    
     ggplot(data =  ma, mapping = aes(x = True, y = Predicted)) +
      labs(title = "Random Forest Confusion Matrix", subtitle = sTitle) +
      geom_tile(aes(fill = freq), colour = "red") +
      geom_text(aes(label = sprintf("%1.0f", freq)), vjust = 1) +
      scale_fill_gradient(low = "lightgreen", high = "green") +
      theme_bw() + theme(legend.position = "none")
      }
rfConfusionMatrix(test2,"RandomForest - 100 Trees Prediction")


#PCA

mushrooms[sapply(mushrooms, is.factor)] <- lapply(mushrooms[sapply(mushrooms, is.factor)], as.numeric) 

str(mushrooms)

mushroomsPCA <- prcomp(mushrooms[-1])
fviz_eig(mushroomsPCA, addlabels=T)
options(repr.plot.height=10)


options(repr.plot.height=30, repr.plot.width=30)
fviz_pca_ind(mushroomsPCA, label="none", habillage=mushrooms$class, 
             addEllipses = TRUE, ellipse.level = 0.95, palette = c("green", "red"))
             
             
options(repr.plot.height=30, repr.plot.width=30)
fviz_pca_var(mushroomsPCA, col.var="cos2",
             gradient.cols=bright, 
             repel=TRUE)


#___________________________________________

##Building model for logistic regression

set.seed(202)
p = runif(nrow(mushrooms), 0, 1)
trainLR <- mushrooms [p < 0.7, ]
testLR <- mushrooms [p > 0.7, ]

str(mushrooms)
modelLR <- glm(class ~ odor + gill.color + ring.type + habitat + 
              stalk.color.below.ring + spore.print.color + population + 
              gill.size + stalk.root + stalk.surface.above.ring + stalk.color.above.ring + 
              stalk.shape + cap.surface + cap.color + stalk.surface.below.ring +
              ring.number + gill.spacing + cap.shape + bruises + veil.color + gill.attachment,                    data = trainLR)
summary(modelLR)

